---
title: "rokf: R Interface to OKF Agent Memory"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rokf: R Interface to OKF Agent Memory}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(rokf)
# Work in a temp bundle so the vignette never touches the reader's files
kb <- file.path(tempdir(), "vignette-kb")
```

## The problem

AI agent conversations are ephemeral. When the context window resets,
architectural decisions, domain discoveries, and operational facts are lost.
OKF Agent Memory persists that knowledge as plain Markdown + YAML inside your
repository (`knowledge/`), so it is Git-tracked, human-readable, and auditable
via `git diff`. Retrieval is local BM25 — no vector database, no embeddings,
no per-query API cost.

`rokf` is the R interface to the `okf` CLI. The package vendors pre-compiled
binaries, so there are no system dependencies to install.

## The workflow: Read — Work — Remember

Every agent session follows one loop: **search before you create**, do the
work, persist what you learned, validate.

### 1. Initialize a bundle

```{r init}
okf_init(kb)
rokf:::is_okf_bundle(kb)
```

### 2. Create concepts (only after searching)

```{r create}
okf_create("decisions/cache-ttl", bundle = kb,
  type  = "Decision",
  title = "Redis Cache TTL",
  desc  = "Default TTL 300s for session data")

okf_create("architecture/sessions", bundle = kb,
  type  = "Architecture",
  title = "Session Store",
  desc  = "Sessions live in Redis behind the API gateway")
```

### 3. Search existing knowledge

```{r search}
okf_search("session TTL", bundle = kb, limit = 5)
```

Search returns a tibble (`rank`, `score`, `id`, `type`, `title`,
`description`, `matches`) — empty when nothing is relevant.

### 4. Read a concept in full

```{r show}
concept <- okf_show("decisions/cache-ttl", bundle = kb)
concept$title
concept$description
```

### 5. Link related concepts

Links are stored as relative Markdown links inside the source concept's body,
so the graph can never fragment into orphans:

```{r relate}
okf_relate("decisions/cache-ttl", "architecture/sessions", bundle = kb,
  desc = "TTL applies to the session store")

# The link is a relative Markdown link in the source concept's body
cat(okf_show("decisions/cache-ttl", bundle = kb)$body)
```

### 6. Validate before you commit

```{r validate}
result <- okf_validate(kb)
result$is_conformant
result$concept_count
```

`okf_validate(kb, strict = TRUE)` additionally gates broken links, orphans,
and provenance gaps as errors — run that in CI or as a pre-commit hook.
Note: the Windows `okf` binary (v0.1.2) fails to resolve relative links, so
strict validation rejects linked bundles on Windows even when the links are
correct. Same-directory links validate strictly on Linux and macOS.

## Agent integrations

For `ellmer` chat agents, `okf_skill()` registers search/show/create/update/
validate as tools. For `tidyllm` pipelines, `okf_verbs()` provides
search/show/persist/validate verbs that inject results as system messages.
Both follow the same search-before-create contract enforced above.
